Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys

### 2023-07-07 analysis/analysis.qmd
(1) area of wedges vs tiles (2) effective number of responses for rounds 6 and 7

neff1 <- n1/d1
neff2 <- n2/d2

@@

lambda <- n1/(d1*denom) 
```

@@

```{r}
#| out-width: 80%
#| fig-cap: "Certainty by task and correctness. Correct answers in the unaligned task have the least certainty associated with them. An incoorect response on the unaligned task shows a significant oost in certainty. "
round_12_certainty_long %>% 
  ggplot(aes(x = Response, weight=weight)) + geom_bar() + 
  ylab("Number of (weighted) responses") + 
  facet_grid(Correct~Question) +
  geom_vline(
    aes(xintercept = estimate),
    colour = "darkorange",
    data = data.frame(
      coefs_certainty, 
      Correct= gsub(".*(FALSE|TRUE)", "\\1", terms),
      Question = gsub("Question(.*):Correct.*", "\\1", terms))
  )
```

@@

A total of three different charts were shown to (a part) of the panelists from Round 6 and all panelists of Round 7. 
The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.

@@

@@

@@

Note: we do not have certainty for round 7. We also do not have timing by individual comparison.

>>>>

neff1 <- n1/d1 # same as neff(round1_adj$weight)
neff2 <- n2/d2

@@

lambda <- n1/(d1*denom) # same as calculate_lambda(round1_adj$weight, round2_adj$weight)
```

@@

```{r}
#| out-width: 80%
#| fig-cap: "Certainty by task and correctness. Correct answers in the unaligned task have the least certainty associated with them. An incoorect response on the unaligned task shows a significant boost in certainty. "
round_12_certainty_long %>% 
  ggplot(aes(x = Response, weight=weight)) + geom_bar() + 
  ylab("Number of (weighted) responses") + 
  facet_grid(Correct~Question) +
  geom_vline(
    aes(xintercept = estimate),
    colour = "darkorange",
    data = data.frame(
      coefs_certainty, 
      Correct= gsub(".*(FALSE|TRUE)", "\\1", terms),
      Question = gsub("Question(.*):Correct.*", "\\1", terms))
  )
```

@@

The data this section is based on comes from (a part) of Round 6 and all responses from Round 7.

@@

A total of three different charts were shown to (a part) of the panelists from Round 6 and all panelists of Round 7. 
Each of the charts has four labelled pieces A, B, C, and D.

The questions asked for each of these charts were of the form: "which of the marked pieces is bigger? Just your best guess is fine".

A.	A or B?

RESPONSE OPTIONS:

1.	A is bigger
2.	B is bigger
3.	They are the same

Each participant was asked a total of six comparisons of this type: A vs B, A vs C, A vs D, B vs C, B vs C, and C vs D. The number of valid responses and the resulting effective sample sizes are shown in @tbl-combine-67

```{r tbl-combine-67}
#| label: tbl-combine-67
#| tbl-cap: Table of the number of effective responses by task 
dframe <- data.frame(
  round = c("Round 6", "Round 7", "Round 7"),
  task = c("Bar", "Pie-Aligned", "Pie-Unaligned"),
  responses = c(nrow(response_6), nrow(response_7 %>% filter(task=="Pie\nA-")), nrow(response_7 %>% filter(task=="Pie\nUn-"))),
  effective = c(neff(response_6$weight), neff((response_7 %>% filter(task=="Pie\nA-"))$weight),
                neff((response_7 %>% filter(task=="Pie\nUn-"))$weight)),
  participants = c(length(unique(response_6$CaseId)),
                   nrow(unique(response_7 %>% filter(task=="Pie\nA-") %>% dplyr::select(CaseId))),
                   nrow(unique(response_7 %>% filter(task=="Pie\nUn-") %>% dplyr::select(CaseId)))),
  lambda = c(calculate_lambda(response_6$weight, response_7$weight), 
             calculate_lambda((response_7 %>% filter(task=="Pie\nA-"))$weight, 
                               c(response_6$weight,(response_7 %>% filter(task=="Pie\nUn-"))$weight)),
             calculate_lambda((response_7 %>% filter(task=="Pie\nUn-"))$weight, 
                               c(response_6$weight,(response_7 %>% filter(task=="Pie\nA-"))$weight)))
)

dframe %>% 
  group_by(round) %>%
  gt() %>%
  tab_header("Number of (effective) responses by task") %>%
  cols_label(
    responses = md("number of responses:<br>nominal"),
    effective = md("number of responses:<br>effective"),
    participants = md("number of participants"),
    lambda = md("λ"),
  ) %>%
  fmt_number(
    everything(), 
    decimals = 0,
    use_seps=TRUE
  ) %>%
  fmt_number(
    columns = effective,
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(lambda),
    decimals = 4
  ) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold"
  )
```

@@

Besides intrinsic reference lines, the overall size of the stimulus might be another factor contributing to the difference in accuracy between the tiles and the wedges. In the shown example, the marked tiles have an  area between $133 \times 55 = 7315$  pixels (for D) to  $144 \times 55 = 7920$ pixels (for B), while the pie wedges  are based on circles with a diameter of 105 pixels, and therefore cover visually more than twice the area of the tiles in the barchart. The area of D in form of a pie-wedge is $\frac{44.4}{100} \times 105^2 \pi  \approx$ 15,378 pixels and B has an area of $\frac{47.6}{100} \times 105^2 \pi  \approx$ 16,487 pixels.

@@

Note: we do not have certainty for round 7. We also do not have timing by individual comparisons.
